---
title: "Basic R"
author: "Dav Clark"
date: "September 19, 2017"
output: html_document
---

```{r, include=FALSE}
library(tidyverse)
```

## Issue #9

*below is a lightly-edited copy-paste of Yusuke's question:

Here, I was trying to make a new data frame (or is this a vector?) with all the mean values of each question:

```{r}
red_day1 <- data.frame(matrix(sample(1, 18, replace=TRUE), ncol=18))
red_day1_mean <- colMeans(red_day1)
red_day1_mean
```

I get the above table with values of only “1”.


I tried this alternative approach:

```{r}
red_day1_mean <- colMeans(day1, na.rm = FALSE, dims = 1:18)
red_day1_mean
```

Dav gets the above error. Yusuke probably meant:

```{r}
red_day1_mean <- colMeans(red_day1, na.rm = FALSE, dims = 1:18)
red_day1_mean
```

In which case we are getting the same error:

```
first element used of 'length.out' argumentError in dimnames(x)[[dims + 1L]] : recursive indexing failed at level 2
```

Any idea why I get these outputs? Any suggestions how to solve this? 

### Dav's response

What's going on with that data frame?

```{r}
red_day1
```

OK - so that is 1 row, 18 columns. Taking the mean of each column is taking a mean of one number. So, you just get the same things back. My guess is you are trying to create a sample data frame that would make sense to take samples over.

To start, let's look at the help for sample (you need to do this on your own, dear reader!). A quote: "sample takes a sample of the specified size from the elements of x using either with or without replacement." Examples:

```{r}
# We MUST use replace=TRUE here, as we are taking more items than choices
# (We can't take 10 unique values from 3 items in `c(1,2,3)`)
sample(c(1,2,3), 10, replace=TRUE)
sample(c(2,3,4), 20, replace=TRUE)
```

So I guess we are trying to make a data frame with some random data. We can feed the above into a data frame constructor. Remember, we are leaning towards tidyverse, which tries to make it harder for us to make mistakes, so we use `data_frame`, not `data.frame` (we can also use the more whimsical and identical `tibble`). There are a couple ways to do this, but here's one of the more sensible ways for now:

```{r}
# This doesn't need to be a ton of typing, as we can copy-paste the sample bit
day1_red_alt <- data_frame(a=sample(1:10, 6, replace=TRUE), 
                           b=sample(1:10, 6),
                           c=sample(1:10, 6))
day1_red_alt
day1_red_alt %>% colMeans()
```

That seems reasonable.

But following closer to what Yusuke initially tried to do:

```{r}
red_day1 <- data.frame(matrix(sample(1:10, 18, replace=TRUE), ncol=3))
red_day1
```

