---
title: "Basic R"
author: "Dav Clark"
date: "September 19, 2017"
output: html_document
---

```{r, include=FALSE}
library(tidyverse)
```

## Issue #9

*below is a lightly-edited copy-paste of Yusuke's question:

Here, I was trying to make a new data frame (or is this a vector?) with all the mean values of each question:

```{r}
red_day1 <- data.frame(matrix(sample(1, 18, replace=TRUE), ncol=18))
red_day1_mean <- colMeans(red_day1)
red_day1_mean
```

I get the above table with values of only “1”.


I tried this alternative approach:

```{r}
red_day1_mean <- colMeans(day1, na.rm = FALSE, dims = 1:18)
red_day1_mean
```

Dav gets the above error. Yusuke probably meant:

```{r}
red_day1_mean <- colMeans(red_day1, na.rm = FALSE, dims = 1:18)
red_day1_mean
```

In which case we are getting the same error:

```
first element used of 'length.out' argumentError in dimnames(x)[[dims + 1L]] : recursive indexing failed at level 2
```

Any idea why I get these outputs? Any suggestions how to solve this? 

### Dav's response

What's going on with that data frame?

```{r}
red_day1
```

OK - so that is 1 row, 18 columns. Taking the mean of each column is taking a mean of one number. So, you just get the same things back. My guess is you are trying to create a sample data frame that would make sense to take samples over.

To start, let's look at the help for sample (you need to do this on your own,
dear reader!). A quote: "sample takes a sample of the specified size from the
elements of x using either with or without replacement." 

But maybe you wanted just ones... this would do that:

```{r}
matrix(1, nrow=3, ncol=6)
```


Examples:

```{r}
# We MUST use replace=TRUE here, as we are taking more items than choices
# (We can't take 10 unique values from 3 items in `c(1,2,3)`)
sample(c(1,2,3), 10, replace=TRUE)
sample(c(2,3,4), 20, replace=TRUE)
```

So I guess we are trying to make a data frame with some random data. We can feed the above into a data frame constructor. Remember, we are leaning towards tidyverse, which tries to make it harder for us to make mistakes, so we use `data_frame`, not `data.frame` (we can also use the more whimsical and identical `tibble`). There are a couple ways to do this, but here's one of the more sensible ways for now:

```{r}
# This doesn't need to be a ton of typing, as we can copy-paste the sample bit
day1_red_alt <- data_frame(a=sample(1:10, 6, replace=TRUE), 
                           b=sample(1:10, 6),
                           c=sample(1:10, 6))
day1_red_alt
day1_red_alt %>% colMeans()
```

That seems reasonable.

But following closer to what Yusuke initially tried to do:

```{r}
red_day1 <- data.frame(matrix(sample(1:10, 18, replace=TRUE), ncol=3))
red_day1
```

## More on #8

The above shows how to create a tibble with column names. You use variable names in the function call, as we use for 'a', 'b' and 'c' above.

But if you forgot to do this ahead of time, you do it like this:

```{r}
colnames(red_day1)
```
```{r}
colnames(red_day1) <- c('new', 'column', 'names')
red_day1
```

# A small note on #7

While I don't think this is what Yusuke was talking about. There are many ways
to format HTML output, though we usually lose the nice paginated table view.
Here, table formatting is handled by the `kable` function in knitr, which is
reasonably flexible. You can see that the caption doesn't look very nice. I'm 
not going to spend the time to make it look nice for now - we'll handle this
kind of thing within flexdashboard.

```{r}
knitr::kable(red_day1, format='html', 
             caption='Some made up data for Team RED, Day 1')
```

# Yusuke was truly working with the ADHD surveys

```{r}
adhd_data <- read_csv('../2017-NZ-reports/deidentified/adhd_survey.csv')
red_day1 <- adhd_data %>% filter(team == 'RED', day == 'day1')
red_day1 %>% select(starts_with('adhd')) %>% colMeans(na.rm=TRUE)
```
```{r}
red_day1 %>% select(starts_with('adhd')) %>% 
  summarise_all(funs(median, mean), na.rm=TRUE)
```

Note: summarize and summary are quite different.
  
```{r}
summary(red_day1)
```


```{r}
simple_df <- data_frame(a=c(1,2), b=c(3,NA))
simple_df
simple_df %>% colMeans(na.rm=TRUE)
```

# To look at student mean across questions

```{r}
colnames(red_day1)
red_day1['recomputed_total'] <- 
  red_day1 %>% select(starts_with('adhd')) %>% rowSums()
red_day1
```

```{r}
red_day1 %>% arrange(attention_total)
```

